Equalizer: A Mature Parallel Rendering Framework

### 2017-04-03 paperIEEE.tex
VR and Equalizer content
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@8 db7c848b-7755-4df2-b763-8796797afa52

\usepackage{color}
\definecolor{RED}{rgb}{1,0,0}
\definecolor{BLUE}{rgb}{0,0,1}
\newcommand{\FIXME}[1]{\textbf{\color{BLUE}{FIXME: #1}}}
\newcommand{\sref}[1]{Section~\ref{#1}}
%\newcommand{\FIXME}[1]{}

@@

\cite{EMP:09} summarized the work in parallel rendering at the time of
publication. In the following we will present the related work published since
this publication.

@@

An observer looks at one or more views. It is described by the observer position
in the world and its eye separation. Each observer will have its own stereo
mode, focus distance and frame loop (framerate). This allows to have untracked
views and multiple tracked views, e.g., two HMDs, in the same application.

@@

Virtual Reality

@@

@@

Focus what user is looking at

@@

Better HMD by measuring user geometry

@@

Constant frame rate for fill-limited applications

@@

Two graphs: Time to render a 4k, 256-step MSAA image of largest ply model and
Livre using 1..16 nodes with all modes (2D, 2D LB, DB, DPlex (over AA steps),
tiles, chunks, pixel, subpixel)

>>>>

\usepackage{color}
\definecolor{RED}{rgb}{1,0,0}
\definecolor{BLUE}{rgb}{0,0,1}
\newcommand{\FIXME}[1]{\textbf{\color{BLUE}{FIXME: #1}}}
\newcommand{\sref}[1]{Section~\ref{#1}}
\newcommand{\fig}[1]{Figure~\ref{#1}}
%\newcommand{\FIXME}[1]{}

@@

In \cite{EMP:09}, we presented the Equalizer parallel rendering framework. This
paper also summarized the work in parallel rendering up to 2009. In the
following we will present the related work published since then.

@@

\label{SEC_observer}
An observer looks at one or more views. It is described by the observer position
in the world and its eye separation. Each observer will have its own stereo
mode, focus distance and frame loop (framerate). This allows to have untracked
views and multiple tracked views, e.g., two HMDs, in the same application.

@@

Virtual Reality is an important field for parallel rendering. It does however
require special attention to support it as a first-class citizen in a generic
parallel rendering framework. Equalizer has been used in many virtual reality
installations, such as the Cave2 (\cite{FNTTL:13}), the high-resolution C6 CAVE
at the KAUST visualization laboratory and head-mounted displays
(\fig{FIG_teaser}). In the following we lay out the features needed support
these installations. All features presented were motivated by application use
cases and have been validated with their respective users.

@@

Head tracking is the minimal feature needed to support immersive
installations. Equalizer does support multiple, independent tracked views
through the observer abstraction introduced in \sref{SEC_observer}. Built-in
VRPN support enables the direct, application-transparent configuration of a VRPN
tracker device. Alternatively, an application can provide a $4\times 4$ tracking
matrix defining the transformation from the canvas coordinate system to the
observer. Both CAVE-like tracking with fixed projection surfaces and HMD
tracking modes are implemented.

@@

To our knowledge, all parallel rendering systems have the focal plane coincide
with the physical display surface. For better viewing comfort, we introduce a
new dynamic focus mode, where the application defines the distance of the focal
plane from the observer, based on the current view direction af the
user. Initial experiments show that this is particularly effective for objects
placed within the immersive space, that is, in front of all display segments.

@@

Traditional head tracking computes the left and right eye positions by using a
configurable interocular distance and the tracking matrix. However, human heads
are not symmetric, and by measuring individual users a more precise frustum can
be computed. Equalizer supports this through the optional configuration (file or
programmatically) of individual 3D positions for the left and right eye.

@@

The DFR equalizer provides a functionality similar to dynamic video resizing
\cite{MBDM:97}, that is, it maintains a constant framerate by adapting the
rendering resolution of a fill-limited application. In Equalizer, this works by
rendering into a source channel (typically on a FBO) separate to the destination
channel, and then scaling the rendering during the transfer (typically through
an on-GPU texture) to the destination channel. The DFR equalizer monitors the
rendering performance and accordingly adapts the resolution of the source
channel and zoom factor for the source to destination transfer. If the
performance and source channel resolutions allows, this will not only subsample,
but also supersample the destination channel to reduce aliasing artefacts.

@@

Two graphs: Time to render a 4k, 256-step MSAA image of largest ply model and
Livre using 1..16 nodes with all modes (2D, 2D LB, DB, DPlex (over AA steps),
tiles, chunks, pixel, subpixel) (+DFR for Livre)
